---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishdatabuilder

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fishdatabuilder)](https://CRAN.R-project.org/package=fishdatabuilder)
<!-- badges: end -->

The goal of fishdatabuilder is to extract the fish monitoring data collected
within the European Water Framework Directive (WFD).

## Installation


<!--You can install the released version of fishdatabuilder from [CRAN](https://CRAN.R-project.org) with:-->

<!--``` r-->
<!--install.packages("fishdatabuilder")-->
<!--```-->

<!--And the development version from [GitHub](https://github.com/) with:-->

You can install the released version of fishdatabuilder from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Fellows-of-the-Fish-Food-Webs-F3W/fishdatabuilder")
```

## Starting

We can extract the french fish monitoring data up to 2022. Those data have
been archived on [Zenodo](https://zenodo.org/records/8099409) at the occasion
of the publication of the [aspe](https://github.com/PascalIrz/aspe) R package
used to manipulate those data.

The first thing to do with the package is to download the zip database and to
extract it:

```{r example, eval = FALSE}
library(fishdatabuilder)
# Download the data
download_optional_data()
# Extract the data
extract_optional_data()
```


The data are located in the cache folder of the package,
using the
[`pkgfilecache`](https://cran.r-project.org/web/packages/pkgfilecache/) R
package.

Now that you did that, you should be able to list and load the files in the
cache of the package:

In development mode:

```{r dev-mode}
library(devtools)
library(pkgfilecache)
load_all()
document()
```


```{r file-list}
head(list_optional_data())
```

You can get station data, clean them, and convert all of them in the same coordinates:

```{r}
station <- clean_station_aspe(
  station = get_raw_station_aspe(),
  ref_coordinates = get_raw_ref_coordinates_station_aspe(),
  crs_to = 4326
  )
head(station)
```



```{r}
#| id: list-cache-files
# Cleaned operation files:
op <- clean_operation_aspe()
head(op)

# There is several operation, to see why
unique(op$operation_id) |> length()
nrow(op)

# Cleaned description of operation files:
op_description <- clean_description_operation_aspe()
head(op_description)
#

```

