---
title: "Getting started"
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: setup
library(fishdatabuilder)
library(magrittr)
```

We can extract the french fish monitoring data up to 2022. Those data have
been archived on [Zenodo](https://zenodo.org/records/8099409) at the occasion
of the publication of the [aspe](https://github.com/PascalIrz/aspe) R package
used to manipulate those data.

The first thing to do with the package is to download the zip database and to
extract it:

```{r example, eval = FALSE}
# Download the data
download_optional_data()
# Extract the data
extract_optional_data()
```


The data are located in the cache folder of the package,
using the
[`pkgfilecache`](https://cran.r-project.org/web/packages/pkgfilecache/) R
package.

Now that you did that, you should be able to list and load the files in the
cache of the package:



```{r file-list}
head(list_optional_files())
```

You can get station data, clean them, and convert all of them in the same coordinates:

```{r}
station <- clean_station_aspe(
  station = get_raw_station_aspe(),
  ref_coordinates = get_raw_ref_coordinates_station_aspe(),
  crs_to = 4326
  )
head(station)
```

## Access cleaned sampling objects

```{r}
#| id: list-cache-files
# Cleaned operation files:
op <- clean_operation_aspe()
head(op)

# Cleaned description of operation files:
op_description <- clean_description_operation_aspe()
head(op_description)

# Cleaned elementary sampling
elementary_sampling <- cleaning_elementary_sampling(
  sampling = get_elementary_sampling_aspe(),
  ref_sampling = get_ref_elementary_sampling_aspe(),
  ref_passage = get_ref_passage_aspe()
)
head(elementary_sampling)

# Cleaning point group
point_group <- cleaning_point_group(
  point_group = get_point_group_aspe(),
  ref_point_group = get_ref_point_group_aspe()
)
head(point_group)

# Cleaned fish batches (lot_poisson in french)
fish_batch <- clean_fish_batch()
head(fish_batch)

# Cleaned individual measurement of fish from the batches
ind_measure <- clean_individual_measurement_aspe()
head(ind_measure)
```

## Filter sampling events and subsampling 

You can filtered the dataset according to some criteria, by default that we believe relevant for spatio-temporal analysis.

```r
filtered_sampling <- filter_operation_batch_measure(
  operation = op,
  op_protocol_to_keep = c("complete", "partial_by_point",  "partial_over_bank"),
  op_objective_to_exclude = vec_op_objective_to_exclude(),
  oldest_sampling_date = "1995-01-01",
  omit_na_site = TRUE,
  point_group = point_group,
  min_prop_point_group_on_bank = .999,
  ele_sampling = elementary_sampling,
  max_passage_number = 1,
  fish_batch = fish_batch,
  ind_measure = ind_measure
)
names(filtered_sampling)
head(filtered_sampling$operation)
head(filtered_sampling$fish_batch)
head(filtered_sampling$ind_measure)
```

## Generate fish individual size

We sample sampling operations because it takes almost 50 minutes using parallelisation.

```r
operation_id_sample <- sample(op$operation_id, 100)
batch_test <- fish_batch |>
  filter(operation_id %in% operation_id_sample) |>
  select(-c(
    weight, estimated_weight, mep_id, tlo_id,
    prelevement_id, operation_id
  ))
measure_test <- ind_measure |>
  filter(operation_id %in% operation_id_sample) |>
  select(-c(site_id:prelevement_id, measure_id))
```

Generate the individual body size from batches and individual measurements.

```r
options(mc.cores = 1)
system.time({
test <- get_size_from_batch(
      batch = batch_test,
      id_var = batch_id,
      batch_type_var = batch_type,
      nb_var = number,
      min_var = min_length,
      max_var = max_length,
      species = species_code,
      measure = measure_test,
      measure_id_var = batch_id,
      size_var = size)
})

test <- test |>
  unnest(cols = size_mm)
nrow(test)
test |>
  head
```

