---
title: "Getting started"
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: setup
library(fishdatabuilder)
library(magrittr)
```

We can extract the french fish monitoring data up to 2022. Those data have
been archived on [Zenodo](https://zenodo.org/records/8099409) at the occasion
of the publication of the [aspe](https://github.com/PascalIrz/aspe) R package
used to manipulate those data.

The first thing to do with the package is to download the zip database and to
extract it:

```{r example, eval = FALSE}
# Download the data
download_optional_data()
# Extract the data
extract_optional_data()
```


The data are located in the cache folder of the package,
using the
[`pkgfilecache`](https://cran.r-project.org/web/packages/pkgfilecache/) R
package.

Now that you did that, you should be able to list and load the files in the
cache of the package:



```{r file-list}
head(list_optional_files())
```

You can get station data, clean them, and convert all of them in the same coordinates:

```{r}
station <- clean_station_aspe(
  station = get_raw_station_aspe(),
  ref_coordinates = get_raw_ref_coordinates_station_aspe(),
  crs_to = 4326
  )
head(station)
```

## Access cleaned sampling objects

```{r}
#| id: list-cache-files
# Cleaned operation files:
op <- clean_operation_aspe()
head(op)

# Cleaned description of operation files:
op_description <- clean_description_operation_aspe()
head(op_description)

# Cleaned elementary sampling
elementary_sampling <- cleaning_elementary_sampling()
head(elementary_sampling)

# Cleaning point group
point_group <- cleaning_point_group()
head(point_group)

# Cleaned fish batches (lot_poisson in french)
fish_batch <- clean_fish_batch()
head(fish_batch)

# Cleaned individual measurement of fish from the batches
ind_measure <- clean_individual_measurement_aspe()
head(ind_measure)
```

## Filter sampling events and subsampling 

You can filtered the dataset according to some criteria, by default that we believe relevant for spatio-temporal analysis.

```r
filtered_sampling <- filter_operation_batch_measure(
  operation = op,
  op_protocol_to_keep = c("complete", "partial_by_point",  "partial_over_bank"),
  op_objective_to_exclude = vec_op_objective_to_exclude(),
  oldest_sampling_date = "1995-01-01",
  omit_na_site = TRUE,
  point_group = point_group,
  min_prop_point_group_on_bank = .999,
  ele_sampling = elementary_sampling,
  max_passage_number = 1,
  fish_batch = fish_batch,
  ind_measure = ind_measure
)
names(filtered_sampling)
head(filtered_sampling$operation)
head(filtered_sampling$fish_batch)
head(filtered_sampling$ind_measure)
```

## Sanitize species code


```{r}
filtered_sampling[c("fish_batch", "ind_measure")] <-
  filtered_sampling[c("fish_batch", "ind_measure")] |>
    purrr::map(\(x) sanitize_species_code(
      data = x,
      species_var = species_code,
      #species_to_remove = species_code_to_remove(),
      #species_to_replace = species_code_to_replace(),
      species_ref = species
    ))
```

## Sanitize fish size measurements

```{r}
nb_ind_fork <- filtered_sampling$fish_batch |>
  filter(length_type == "fork") %>%
  summarise(n = sum(number)) |> pull(n)
nb_ind_fork / sum(filtered_sampling$fish_batch$number)
```

```{r}
#| id: length-conversion
species_fork_length <- filtered_sampling$fish_batch |>
    filter(length_type == "fork") |>
    distinct(species_code) |>
    left_join(species,
      by = join_by(species_code)
    )
ll <- rfishbase::length_length(species_list = species_fork_length$latin_name)

sanitized_length_type <- convert_fork_to_total(
  fish_batch = filtered_sampling$fish_batch,
  ind_measure = filtered_sampling$ind_measure,
  species_ref = cleaning_species_ref_aspe(),
  fishbase_length_length = ll,
  conversion_vector = coefficients_fork2total(),
  convert_intercept_cm2mm = TRUE,
  manual_priority = FALSE,
  verbose = TRUE
)
sanitized_length <- remove_impossible_lengths(
  ind_measure = sanitized_length_type$ind_measure,
  fish_batch = sanitized_length_type$fish_batch,
  species_ref = species,
  remove_outliers = TRUE
)
```

```{r}
sanitized <- sanitize_batch_data(
  fish_batch = sanitized_length$fish_batch,
  ind_measure = sanitized_length$ind_measure,
  species_ref = species,
  min_individuals_G = 2,
  min_individuals_SL = 6
)
```

## Generate fish size from batch

```{r}
ti <- generate_individual_sizes_vectorized(sanitized, verbose = TRUE)
```
